---
title: "Political Control w/ Cleaned data"
output: html_document
date: "2025-03-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Do Republican-controlled states ban more books than Democrat-controlled states

```{r echo=TRUE}
library(ggplot2)
library(tidyverse)
library(dplyr)
new_data <- read_csv("cleaned_data.csv")

leg_control_counts <- new_data %>% 
  group_by(Year, State, Leg.Control) %>% 
  summarise(Count = n(), .groups = 'drop') %>% 
    mutate(Leg.Control = recode(Leg.Control,
                              "Dem" = "Democratic",
                              "Rep" = "Republican"))

leg_control_counts

ggplot(leg_control_counts, aes(x = factor(Year), y = Count, fill = Leg.Control)) +
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(
    values = c(
      "Democratic" = "blue",
      "Republican" = "red",
      "Divided" = "purple",
      "N/A" = "gray"
    )
  ) + 
  labs(
    title = "Do State's with Republican Controlled Legislatures Banning More Books?", 
    subtitle = "It would appear so", 
    caption = "N/A refers to Nebraska, which has a nonpartisan unicameral legislature",
    x = "Year",
    y = "Banned Books Count", 
    fill = "Party in Legislative control"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle= 45, hjust =1))

```

# State by State - faceted

```{r echo=TRUE, fig.height=20, fig.width=20}



# Faceted bar plot by state
ggplot(leg_control_counts, aes(x = factor(Year), y = Count, fill = Leg.Control)) +
  geom_bar(stat = "identity", position = "dodge") + 
  facet_wrap(~ State, scales = "fixed") +  # Facet by state
  scale_fill_manual(
    values = c(
      "Democratic" = "blue",
      "Republican" = "red",
      "Divided" = "purple",
      "N/A" = "gray"
    )
  ) + 
  labs(
    title = "Now Lets Look at Each State! Do States with Republican Controlled Legislatures Ban More Books?",
    subtitle = "Overwhemlingly so",
    caption = "N/A refers to Nebraska, which has a nonpartisan unicameral legislature",
    x = "Year",
    y = "Banned Books Count",
    fill = "Party in Legislative Control"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 8)  # Adjust strip size for readability
  )

```

# Governor's Party

```{r echo=TRUE}
gov_party_counts <- new_data %>% 
  group_by(Year, State, Gov.Party) %>% 
  summarise(Count = n(), .groups = 'drop') %>%
    mutate(Gov.Party = recode(Gov.Party,
                              "Dem" = "Democrat",
                              "Rep" = "Republican"))

ggplot(gov_party_counts, aes(x = factor(Year), y = Count, fill = Gov.Party)) +
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(
    values = c(
      "Democrat" = "blue",
      "Republican" = "red",
      "Divided" = "purple",  # in case it appears
      "N/A" = "gray"
    )
  ) + 
  labs(
    title = "Do States with Republican Governors Ban More Books?",
    subtitle = "It would appear so",
    x = "Year",
    y = "Banned Books Count",
    fill = "Governor's Party"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

# Top 10 States with most banned books - Incorporating State Political Leanings

```{r echo=TRUE}
top_states <- new_data %>%
  group_by(State) %>%
  summarise(Total_Bans = n(), .groups = "drop") %>%
  arrange(desc(Total_Bans)) %>%
  slice_head(n = 10) %>% 
  pull(State)


top_states_by_year <- new_data %>% 
  filter(State %in% top_states) %>%
  group_by(Year, State, Leg.Control, Gov.Party) %>%
  summarise(X = n(), .groups = "drop") %>%
  mutate(
    Leg.Control = recode(Leg.Control,
                         "Dem" = "Democratic",
                         "Rep" = "Republican")
  ) %>% 
  arrange(desc(X))

top_states_by_year
  
```

# Visualization with faceted heat map

```{r echo=TRUE}
ggplot(top_states_by_year, aes(x = factor(Year), y = reorder(State, desc(State)), fill = X)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "orange", high = "red") +
  facet_wrap(~Leg.Control) + 
  labs(
    title = "10 States with most Banned Books by Year and Legislative Control",
    subtitle = "Darker tiles indicate more bans",
    x = "Year",
    y = "State",
    fill = "Banned Books Count",
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 10, face = "bold"),
    panel.grid = element_blank()
  )
```

# Origin of Challenge

From <https://pen.org/report/banned-usa-growing-movement-to-censor-books-in-schools/#heading-7> Legislative efforts to pass educational "gag orders"; some state legislators proposing or supporting bills that impact the selection and removal of books in school classrooms and libraries

Flordia's Parental Rights in Education law signed March 2022 - referred to as the "Don't Say Gay" law

Georgia SB26 passed March 22 makes it easier to remove books with "offensive content"

Tennessee SB2247 makes it easier for books to be banned form studnet access statewide on the basis of challenges filed in individual districts

Utah HB 374 Sensitive Materials in Schools signed into law March 2022 - prohibits sensitive instructura materials considered pornographic or indecent.

Missouri SB 775 - amendment that makes it Class A misdemeanor if person “affiliated with a public or private elementary or secondary school” provides “explicit sexual material” to a student. St. Louis area school districts reacted preemptively by remobing books, especially graphic novels

```{r echo=TRUE}
origin_trends <- new_data %>%
  filter(!is.na(Origin.of.Challenge)) %>%
  group_by(Year, Origin.of.Challenge) %>%
  summarise(X = n(), .groups = "drop") %>%
  arrange(desc(X))

origin_trends
```

```{r echo=TRUE}
ggplot(origin_trends, aes(x = factor(Year), y = X, color = Origin.of.Challenge, group = Origin.of.Challenge)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
   scale_color_manual(
    values = c(
      "Administration" = "#d73027",
      "Formal Challenge" = "#4575b4",
      "Administrator" = "#fdae61",
      "Informal Challenge" = "#66bd63",
      "Informal Challenge, Administration" = "#fee08b",
      "Administration/Formal Challenge" = "green",
      "Legislative Action" = "#9e0142",
      "Administration/Informal Challnge" = "grey",
      "Other"= "lightblue",
      "Unclear" = "purple",
      "Formal Challenge, Administration" = "pink"
    ),
    na.translate = FALSE  # avoid NA color showing up if not desired
  ) +
  labs(
    title = "Trends in Book Ban Challenges by Origin",
    subtitle = "Who is initiating the bans?",
    x = "Year",
    y = "Number of Banned Book Entries",
    color = "Origin of Challenge"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )

```

```{r echo=TRUE}
origin_trends_focus <- new_data %>%
  filter(!is.na(Origin.of.Challenge)) %>%
  mutate(
    Challenge_Type = case_when(
      grepl("Administration|Administrator", Origin.of.Challenge, ignore.case = TRUE) ~ "Administrative",
      grepl("Legislative|Legislature", Origin.of.Challenge, ignore.case = TRUE) ~ "Legislative",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Challenge_Type)) %>%
  group_by(Year, Challenge_Type) %>%
  summarise(Count = n(), .groups = "drop")

origin_trends_focus
```

```{r echo=TRUE}
ggplot(origin_trends_focus, aes(x = factor(Year), y = Count, color = Challenge_Type, group = Challenge_Type)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
    title = "Trends in Book Ban Challenges: Administration vs. Legislative",
    subtitle = "Yearly counts of bans initiated by administrative or legislative action",
    x = "Year",
    y = "Number of Banned Book Entries",
    color = "Origin of Challenge"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```

# What states had legislative action as origin of challenge?

```{r echo=TRUE}


# Filter for legislative origin of challenge
legislativeaction_states <- new_data %>%
  filter(grepl("Legislative", Origin.of.Challenge, ignore.case = TRUE)) %>%
  group_by(State) %>%
  summarise(Count = n(), .groups = "drop") %>%
  arrange(desc(Count))

legislativeaction_states
# Bar plot of states with legislative-origin bans
ggplot(legislativeaction_states, aes(x = reorder(State, Count), y = Count)) +
  geom_col(fill = "#9e0142") +
  coord_flip() +
  labs(
    title = "States with Book Bans Originating from Legislative Action",
    subtitle = "Filtered by entries that include 'legislative' in origin",
    caption = " Note that Missouri, Florida, Utah, Viriginia had Republican Legislatures in 2022",
    x = "State",
    y = "Number of Banned Book Entries",
  ) +
  theme_minimal()

```

# Did increased legislative action in 2022 lead to a surge of administrative challenges the following year?

```{r echo=TRUE}
admin_states_2023 <- new_data %>%
  filter(Year == 2023,
         grepl("Administration|Administrator", Origin.of.Challenge, ignore.case = TRUE)) %>%
  group_by(State) %>%
  summarise(Count = n(), .groups = "drop") %>%
  arrange(desc(Count))

# Plot
ggplot(admin_states_2023, aes(x = reorder(State, Count), y = Count)) +
  geom_col(fill = "#d73027") +
  coord_flip() +
  labs(
    title = "States with Administrative-Origin Book Bans in 2023",
    x = "State",
    y = "Number of Banned Book Entries",
    subtitle = "Filtered by 'Administration' or 'Administrator' in origin field",
  ) +
  theme_minimal()
```

```{r echo=TRUE}

# Extract counts by Year and Origin Type
leg_admin_trends <- new_data %>%
  filter(!is.na(Origin.of.Challenge)) %>%
  mutate(
    Challenge_Type = case_when(
      grepl("Legislative", Origin.of.Challenge, ignore.case = TRUE) ~ "Legislative",
      grepl("Administration|Administrator", Origin.of.Challenge, ignore.case = TRUE) ~ "Administrative",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Challenge_Type)) %>%
  group_by(Year, Challenge_Type) %>%
  summarise(Count = n(), .groups = "drop") %>%
  filter(Year %in% c(2021, 2022, 2023))  # focus on key years

# Plot
ggplot(leg_admin_trends, aes(x = factor(Year), y = Count, fill = Challenge_Type)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("Legislative" = "#9e0142", "Administrative" = "#d73027")) +
  labs(
    title = "Did 2022 Legislative Action Lead to 2023 Administrative Challenges?",
    subtitle = "Comparing counts of legislative and administrative-origin book bans by year, it would appear so!",
    x = "Year",
    y = "Number of Banned Book Entries",
    fill = "Challenge Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
